---
outputs:
- Reveal
title: Develop packages together
hidden: true
layout: list
weight: 11
output: hugodown::md_document
countdown: true
---

# Develop packages together

```{r, echo=FALSE}
library("magrittr")
get_pic <- function(id, alt, width) {
  
  info <- httr::GET(
    paste0("https://api.pexels.com/v1/photos/", id),
    httr::add_headers(Authorization = Sys.getenv("PEXELS_KEY"))
  ) %>%
    httr::content()
  download.file(info$src$large, paste0(id, ".jpeg"))
  cat("```{=html}\n")
cat(paste0('{', '{< figure src="',
paste0(id, ".jpeg"),
'" alt="',
alt,
'" caption="Picture by [',
info$photographer,
' on Pexels](',
info$url,
')." width="',
width , '" >}', '}'))
cat("\n```\n")
}
```

```{r red, results="asis", cache=TRUE, echo=FALSE}
get_pic(id = 7105805, alt = "Plastic rocket ship", 200)
```


---

## rOpenSci software review

Non adversarial, constructive.

Helps disseminate best practice.

Builds a community of practice.

---

## rOpenSci software review

- Reviewer guide https://devguide.ropensci.org/reviewerguide.html
- Review template https://devguide.ropensci.org/reviewtemplate.html
- Packaging guide https://devguide.ropensci.org/building.html

---

## rOpenSci software review

Read examples at https://github.com/ropensci/software-review/issues/ for instance for a package you know (via https://ropensci.org/packages/all/)

---

## Pull request review

- Smaller scope.

- Helps strengthen the code, comments, docs, tests.

- Helps share knowledge.

---

## Pull request review tips

- [usethis helpers](https://usethis.r-lib.org/articles/pr-functions.html) (or GitHub Desktop)?

- use line comments including code suggestions.

- https://davisvaughan.github.io/code-review/

Let's demo this live on the [course repo](https://github.com/maelle/2023-pkg-dev-rocket).


---

## Review each other's work!

In breakout rooms, exchange PRs to read, comment on. 

Ask questions, make a few suggestions.

Be nice :smile_cat:

---

## Debriefing

Anything to report from the breakout rooms?

---

## Conclusion

- collection of tips/patterns;
- importance of team work.

---

## How to keep learning?

- Read blogs,
- Follow the development of your favorite packages,
- Review packages and PRs + have yours reviewed.